---
title: "Lab: Week 2"
author: "Your Name Here"
format: 
  html:
    self-contained: true
execute:
  echo: true
editor: source
---

Monday's lecture: [Chapter 1](https://github.com/samanthaseals/STA4222/blob/master/slides/L01.qmd)

Wednesday's lecture: [Chapter 2](https://github.com/samanthaseals/STA4222/blob/master/slides/L02.qmd)

1. Call in the necessary packages:

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(SDAResources)
```

2. Call in the **agpop** dataset from the [`SDAResources` package](https://rdrr.io/cran/SDAResources/man/agpop.html).

```{r}

```

3. Take a simple random sample of $n=300$ farms. Remember to set your seed before drawing the sample.

```{r}

```

4. Construct a variable (columnn) to indicate if the county has more than 50 large farms in 1992 (*largef92*). 

```{r}

```

5. Use the `count()` function to determine how many counties have more than 50 large farms using the variable you created in the previous question.

```{r}

```

6. Estimate the average number of large farms in 1992 (*largef92*) -- construct both the point estimate and the 95% CI.

```{r}

```

7. Estimate the total number of large farms in 1992 (*largef92*) -- construct both the point estimate and the 95% CI.

```{r, echo = TRUE}

```

8. Estimate the proportion of counties with more than 50 large farms in 1992 (use the variable you constructed question 4) -- construct both the point estimate and the 95% CI.

```{r}

```

9. Challenge and bonus: are these estimates significantly different from 1982? 

10. Challenge and bonus: construct a data visualization that compares the data from 1982 and 1992.

