---
title: "Lab: Week 3"
author: "Your Name Here"
format: 
  html:
    self-contained: true
execute:
  echo: true
editor: source
---

Monday's lecture: [Chapter 3: Part 1](https://github.com/samanthaseals/STA4222/blob/master/slides/L03.qmd)

- Remember that the examples in class provided us with the stratum-level summary statistics. 

- In today's lab, you will need to use knowledge from Chapter 2 to find the stratum-level summary statistics before constructing the overall estimate under stratified sampling.

1. Call in the necessary packages:

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(SDAResources)
```

2. Call in the **agstrat** and **agpop** datasets from the [`SDAResources` package](https://rdrr.io/cran/SDAResources/man/agpop.html).

```{r}

```

3. Create the dataset of summary statistics needed for estimating the mean and total.

```{r}
summaries <- agstrat %>% 
  group_by(region) %>%
  summarize(mean = mean(largef92), # request mean
            var = var(largef92), # request variance
            n = n()) # request sample size

pop <- agpop %>%
  group_by(region) %>%
  summarize(N = n())

summaries <- full_join(summaries, pop, by = "region")
```

4. Estimate the average number of large farms in 1992 (*largef92*) -- construct both the point estimate and the 95% CI.

```{r}

```

5. Estimate the total number of large farms in 1992 (*largef92*) -- construct both the point estimate and the 95% CI.

```{r, echo = TRUE}

```

6. Return to **agstrat** and construct a variable (column) to indicate if the county has more than 50 large farms in 1992 (*largef92*). 

```{r}

```

7. Determine how many counties, by region, have more than 50 large farms using the variable you created in the previous question.

```{r}

```

8. Create the dataset of summary statistics needed for estimating the proportion. (Hint: use the code from question 3 as your starting point - copy and paste into this code chunk and edit to what you need.)

```{r}

```

9. Estimate the proportion of counties with more than 50 large farms in 1992 (use the variable you constructed question 3) -- construct both the point estimate and the 95% CI.

```{r}

```

10. How do these estimates compare to the estimates from last week? Are our estimators more or less precise? Justify your response.

