---
title: "Lab: Week 8"
author: "Your Name Here"
format: 
  html:
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

Monday's lecture: [Chapter 5: Part 1](https://github.com/samanthaseals/STA4222/blob/master/slides/L07.qmd)

The **ozone** dataset contains hourly ozone readings from a site in Monterey County, California for 2018 and 2019.

```{r}
library(SDAResources)
library(tidyverse)
data(ozone)
ozone <- ozone %>%
  pivot_longer(
    cols = starts_with("hr"),
    names_to = "hour",
    names_prefix = "hr",
    values_to = "reading"
  ) %>%
  mutate(num = row_number())
```

1. Create a histogram of the population ozone readings (*reading*). Hint: [here's an example](https://r-graph-gallery.com/220-basic-ggplot2-histogram.html).

```{r}

```

2. Create a scatterplot and identify the type of population this is. Put reading number (*num*) on the *x*-axis and ozone reading (*reading*) on the *y*-axis. Hint: [here's an example](https://r-graph-gallery.com/272-basic-scatterplot-with-ggplot2.html)

```{r}

```

3. Find the mean and standard deviation of the population readings.

```{r}

```

4. The following function will allow us to take systematic samples,

```{r}
getsys = function(k,N){
  n = ceiling(N/k)
  r = sample(1:k, 1)
  seq(r, r + k*(n-1), k)
}

# example use - penguin dataset with k = 5
# sample <- palmerpenguins::penguins[getsys(k = 5, N = nrow(palmerpenguins::penguins)), ]
```

4a. Take a systematic sample with $k = 5$ (i.e., every 5th observation). Estimate the mean - make sure you include the point estimate and the 95% CI.

```{r}

```

4b. Take a systematic sample with $k = 8$ (i.e., every 8th observation). Estimate the mean - make sure you include the point estimate and the 95% CI.

```{r}

```

4c. Take a systematic sample with $k = 12$ (i.e., every 12th observation). Estimate the mean - make sure you include the point estimate and the 95% CI.

```{r}

```

4d. Take a systematic sample with $k = 24$ (i.e., every 24th observation). Estimate the mean - make sure you include the point estimate and the 95% CI.

```{r}

```

4e. Take a systematic sample with $k = 48$ (i.e., every 48th observation). Estimate the mean - make sure you include the point estimate and the 95% CI.

```{r}

```

4f. Take a systematic sample with $k = 50$ (i.e., every 50th observation). Estimate the mean - make sure you include the point estimate and the 95% CI.

```{r}

```

4g. Fill in the table of results:

<center>
|$k$|$\bar{y}_{\text{sys}} \text{ (95\% CI)}$|
|-|:-:|
|5| FILL IN HERE |
|8| FILL IN HERE |
|12| FILL IN HERE |
|24| FILL IN HERE |
|48| FILL IN HERE |
|50| FILL IN HERE |
</center>

5. Challenge (must be completed for full credit)! Create a scatterplot to graphically display the results. Hint 1: use the table in 4g to create a dataset to use for graphing. Hint 2: [here's how to include error bars](https://r-graph-gallery.com/4-barplot-with-error-bar.html).

```{r}

```

6. Write a couple of sentences summarizing what you found in lab. Things to think about -- What happens as we increase the choice of $k$? What happens when we pull from the same time every day (i.e., $k=24$) vs. not (i.e., $k=5$ or $k=50$)? What happens when we pull every couple of days (i.e., $k=48$ or $k=50$) vs. more often ($k<48$)?




