---
title: "Lab: Week 12"
author: "Your Name Here"
format: 
  html:
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

Monday's lecture: [Chapter 6: Part 2](https://github.com/samanthaseals/STA4222/blob/master/slides/L10.qmd)

Consider the **baseball** dataset, which contains statistics on 797 baseball players from the rosters of all major league teams in November 2004. In this lab, we will take a two-stage cluster sample (without replacement) with the teams (*team*) as the clusters and probabilities proportional to the total number of runs (*run*) scored for the teams. 

```{r}
library(tidyverse)
library(SDAResources)
data(baseball)
```

1. Construct the selection probabilities. Remember, teams (*team*) define the clusters and we want probabilities proportional to the total number of runs (*run*) scored. 

```{r}

```

2. Take a sample such that your final sample has approximately 150 players. Write a few sentences to describe (and justify) how you selected your sample.

```{r}

```

3. Use your sample to estimate the mean salary (*salary*) in 2004. Make sure you include the point estimate and the 95% CI.

```{r}

```

4. (Challenge!) Use your sample to estimate the proportion of players that were pitchers (*pos*) in 2004. Make sure you include the point estimate and the 95% CI.

```{r}

```





