---
title: "Lab: Week 14 - part 1"
author: "Your Name Here"
format: 
  html:
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
editor: source
---

Recall the **baseball** dataset, which contains statistics on 797 baseball players from the rosters of all major league teams in November 2004. 

```{r}
library(tidyverse)
library(SDAResources)
data(baseball)
```

1. Determine the appropriate sample size to estimate the mean salary (*salary*) in 2004 using the teams as the strata under optimal allocation and a margin of error of $25,000. 

```{r}

```

2. Draw a stratified sample based on your sample size(s) in question 1.

```{r}

```

3. Estimate the mean salary (*salary*) in 2004 using the sample you just drew. Make sure you include the point estimate and the 95% CI.

```{r}

```

4. How does today's estimate compare to last week's estimate? Why do you think the difference exists?

5. How does today's estimate compare to the population mean? Why do you think the difference exists?






